---
title: "NYPD_prj"
author: "VRG"
date: "February 28, 2022"
output:
  html_document: default
  pdf_document: 
    fig_caption: yes
---

### IMPORT the libraries for the project.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Library
#install.packages("ggplot2") 
# Install & load ggplot2 package


library(dplyr)
library(tidyr)
library("gridExtra")
library(sqldf)
library("ggplot2")

```



## R Markdown
 
This data set has a list of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year.

This is a breakdown of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year. This data is manually extracted every quarter and reviewed by the Office of Management Analysis and Planning before being posted on the NYPD website. Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included. This data can be used by the public to explore the nature of shooting/criminal activity. Please refer to the attached data footnotes for additional information about this dataset.


### Source/ Reference : [https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8]



# Import data and About the data set

```{r , echo=FALSE}   
library(readr)   
DS <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")

```      

```{r, echo=FALSE}
summary(DS)

```

## Preprocessing - Combine Date and change format to Date.

```{r, echo=FALSE}   

## Preprocessing

DS$OCCUR_DATE <- paste(DS$OCCUR_DATE, DS$OCCUR_TIME, sep=" ") # Convert to Date with " "

```    



```{r, echo=FALSE}   
#Tidying and Transforming 

DS[['OCCUR_DATE']] <- strptime(DS[['OCCUR_DATE']], format = "%m/%d/%Y %H:%M:%S")

DS[['OCCUR_DATE_TIME']] <- as.POSIXct(DS$OCCUR_DATE,tz=Sys.timezone())  # Date_Time column

DS$OCCUR_DATE <- as.Date(DS$OCCUR_DATE)

summary(DS)   
head(DS)   

DS[, c('OCCUR_DATE_TIME','OCCUR_DATE','OCCUR_TIME','INCIDENT_KEY' ,'JURISDICTION_CODE', 'LOCATION_DESC','BORO', 'PRECINCT', 'STATISTICAL_MURDER_FLAG','PERP_RACE', 'PERP_AGE_GROUP', 'PERP_SEX', 'VIC_AGE_GROUP', 'VIC_SEX', 'VIC_RACE','Latitude','Longitude','Lon_Lat')]

DS$PERP_RACE[is.na(DS$PERP_RACE)]  <- 'UNKNOWN'

```    

  
```{r, echo=FALSE}    
#Select fetures  

#DS %>% group_by(BORO) %>% summarise(count_incidents = n())

DS_Plot <- DS %>% group_by(BORO) %>% summarise(count_incidents = n())

#View(DS)
head(DS_Plot)
```    


#Select fetures    
```{r, echo=FALSE}    

#install.packages("sqldf")
library(sqldf)

DS_Plot <- DS %>% group_by(BORO,PERP_RACE) %>% summarise(count_incidents = n())

#View(DS)
View(DS_Plot)
```  

#Select fetures    
```{r, echo=FALSE}    

#install.packages("sqldf")


DS_Plot_v <- DS %>% group_by(BORO,VIC_RACE) %>% summarise(count_incidents = n())

#View(DS)
#View(DS_Plot_v)
```



```{r, echo=FALSE}    
#Visualize Data set
# Library
#install.packages("ggplot2")                       # Install & load ggplot2 package


ggplot(DS_Plot_v,                                      # Grouped barplot using ggplot2
       aes(x = BORO,
           y = count_incidents,
           fill = VIC_RACE)) +
  geom_bar(stat = "identity",
           position = "dodge")

```


```{r, echo=FALSE}    
#Visualize Data set

ggplot(DS_Plot,                                      # Grouped barplot using ggplot2
       aes(x = BORO,
           y = count_incidents,
           fill = PERP_RACE)) +
  geom_bar(stat = "identity",
           position = "dodge")

```



```{r, echo=FALSE}    
#Visualize Data set

DS_Plot_t <- DS %>% group_by(OCCUR_DATE) %>% summarise(count_incidents = n())

#View(DS)
#View(DS_Plot_t)


ggplot(DS_Plot_t,                                      # Grouped barplot using ggplot2
       aes(x = OCCUR_DATE,
           y = count_incidents)) +
  geom_bar(stat = "identity",
           position = "dodge")

```




```{r, echo=FALSE}    

#Visualize Data set 
DS$OCCUR_DATE_MONTH <- format(DS$OCCUR_DATE, "%m")

DS_Plot_m <- DS %>% group_by(OCCUR_DATE_MONTH ) %>% summarise(count_incidents = n())

#View(DS)
#View(DS_Plot_m)


ggplot(DS_Plot_m,                                      # Grouped barplot using ggplot2
       aes(x = OCCUR_DATE_MONTH,
           y = count_incidents)) +
  geom_bar(stat = "identity",
           position = "dodge")

```




  
```{r, echo=FALSE}    
#Visualize Data set 

DS$OCCUR_TIME_HOUR <- format(strptime(DS$OCCUR_TIME,"%H:%M:%S"),'%H')

DS_Plot_h <- DS %>% group_by(OCCUR_TIME_HOUR ) %>% summarise(count_incidents = n())

#View(DS)
#View(DS_Plot_m)


ggplot(DS_Plot_h,                                      # Grouped barplot using ggplot2
       aes(x = OCCUR_TIME_HOUR,
           y = count_incidents)) +
  geom_bar(stat = "identity",
           position = "dodge")

```





```{r, echo=FALSE}    
#Analyze Data set

DS$OCCUR_TIME_HOUR <- format(strptime(DS$OCCUR_TIME,"%H:%M:%S"),'%H')

DS_Plot_h_pr <- DS %>% group_by(OCCUR_TIME_HOUR,PERP_RACE ) %>% summarise(count_incidents = n())


DS_Plot_h_pr$PERP_RACE[is.na(DS_Plot_h_pr$PERP_RACE)]  <- 'UNKNOWN'

ggplot(DS_Plot_h_pr, aes(x = OCCUR_TIME_HOUR, y = count_incidents, fill = PERP_RACE)) + geom_bar(stat = "identity")
```

    
```{r, echo=FALSE}    
#Analyze Data set

DS$OCCUR_TIME_HOUR <- format(strptime(DS$OCCUR_TIME,"%H:%M:%S"),'%H')

DS_Plot_h_vic <- DS %>% group_by(OCCUR_TIME_HOUR,VIC_RACE) %>% summarise(count_incidents = n())

ggplot(DS_Plot_h_vic, aes(x = OCCUR_TIME_HOUR, y = count_incidents, fill = VIC_RACE)) + geom_bar(stat = "identity")
```
    
```{r, echo=FALSE}    
#Visualize Data set

Plot1 <- ggplot(DS_Plot_h_pr, aes(x = OCCUR_TIME_HOUR, y = count_incidents, fill = PERP_RACE)) + geom_bar(stat = "identity") +  theme(legend.position = "none")
Plot2 <- ggplot(DS_Plot_h_vic, aes(x = OCCUR_TIME_HOUR, y = count_incidents, fill = VIC_RACE)) + geom_bar(stat = "identity") +  theme(legend.position = "none")
grid.arrange(Plot1, Plot2, ncol=2)

```


```{r, echo=FALSE}    
#Model Data set   

DS_Plot_h_pr$OCCUR_TIME_HOUR <- as.numeric(DS_Plot_h_pr$OCCUR_TIME_HOUR)
DS_Plot_h_vic$OCCUR_TIME_HOUR <- as.numeric(DS_Plot_h_vic$OCCUR_TIME_HOUR)

fit1 = lm(count_incidents~OCCUR_TIME_HOUR, data=DS_Plot_h_vic)
summary(fit1)


```



```{r, echo=FALSE}    
#Visualize Model output

Plot_lm <-ggplot(DS_Plot_h_vic, aes(x=OCCUR_TIME_HOUR, y=count_incidents)) + geom_point()

Plot_lm <- Plot_lm + geom_smooth(method="lm", col="red")

pred_lm = predict(fit1)

Plot_lm <- Plot_lm + geom_line(aes(y = pred_lm), size = 1)

Plot_lm

```   

```{r, echo=FALSE}    
#Select fetures  

#DS %>% group_by(BORO) %>% summarise(count_incidents = n())

DS_map_plot <- DS %>% group_by(BORO) %>% summarise(count_incidents = n(), mean_lat = mean(Latitude,na.rm = TRUE), mean_long = mean(Longitude,na.rm = TRUE) )

DS_map_plot$map_circ_rad <- (DS_map_plot$count_incidents/sum(DS_map_plot$count_incidents))*100

#View(DS)
head(DS_map_plot)
```    
 
 
 
```{r, echo=FALSE}    
#Visualize Model output
library(leaflet)
# create leaflet map
pal <- colorFactor(c("navy", "red", "green","orange","purple"),
                   domain = unique(DS_map_plot$BORO))

DS_ll <- leaflet(DS_map_plot) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~mean_long,
                   lat = ~mean_lat,
                   color = ~pal(BORO),
                   weight = 1, 
                   radius = ~map_circ_rad,
                   stroke = FALSE, fillOpacity = 0.9,
                   label = ~as.character(count_incidents),
                   popup = ~BORO)
DS_ll

```  
